Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-16 paper.tex
wording

To facilitate the adoption of the library, we also try to limit to a bare
minimum the set of external dependencies. As such, NumPy and SciPy -- the two
most common dependencies in the Python scientific ecosystem -- are actually the
only two required dependencies for the library to fully work. Some modules may
rely on other external libraries, but in such cases we tend to prefer to
directly ship them within the codebase rather than adding new dependencies. In
particular, the codebase includes modified versions of the \textsf{LIBSVM} and
\textsf{LIBLINEAR} libraries \citep{chang2011libsvm,fan2008}, both written in
C{}\verb!++! and wrapped using Cython modules.

Finally, some algorithms in \sklearn also support multi-core architectures.
Machine learning algorithms can indeed often be decomposed into independent
smaller tasks that may each be solved in parallel, thereby reducing the overall
computing times. At this point in time, support for multi-core programming in
Python -- at least in the CPython implementation targeted by scikit-learn -- is
however suboptimal. The interpreter supports multithreading, but due to the
\textit{global interpreter lock}, no two threads in a process may execute Python
code at the same time. It follows that multithreading is not a viable solution
for parallelizing parts of an algorithm, even for embarrassingly parallel
problems, such as fitting and evaluating multiple models in a grid search
procedure. The workaround that we use is to rely on the Python multiprocessing
module, which emulates threads using multiple operating system- level processes.
However this solution is not perfect as it may in turn lead to a large memory
overhead since objects (e.g., input datasets) need to be copied within each
process own memory space. In some cases, it may unfortunately even happen that
the overhead becomes so significant that fully exploiting a multi- core
architecture becomes infeasible due memory constraints. As such, a proper
implementation of parallel machine learning algorithms remain a challenging task
for the project.

>>>>

To facilitate the adoption of the library, we also try to limit to a bare
minimum the set of external dependencies. As such, NumPy and SciPy -- the two
most common dependencies in the Python scientific ecosystem -- are actually the
only two required dependencies for the library to fully work. Some modules may
rely on other external libraries, but in such cases we tend to prefer to
directly ship them within the codebase rather than adding new dependencies for
the user to install. In particular, the codebase includes modified versions of
the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using
Cython modules.

Finally, some estimators in \sklearn also support multi-core architectures.
Machine learning algorithms can indeed often be decomposed into independent
smaller tasks that may each be solved in parallel, thereby reducing the overall
computing times. At this point in time, support for multi-core programming in
Python -- at least in the CPython implementation targeted by scikit-learn -- is
however suboptimal. The interpreter supports multithreading, but due to the
\textit{global interpreter lock}, no two threads in a process may execute Python
code at the same time. It follows that multithreading is not a viable solution
for parallelizing parts of an algorithm, even for embarrassingly parallel
problems, such as fitting and evaluating multiple models in a grid search
procedure. The workaround that we use is to rely on the Python multiprocessing
module, which emulates threads using multiple operating system- level processes.
However this solution is not perfect as it may in turn lead to a large memory
overhead since objects (e.g., input datasets) need to be copied within each
process own memory space. In some cases, it may unfortunately even happen that
the overhead becomes so significant that fully exploiting a multi- core
architecture becomes infeasible due memory constraints. As such, a proper
implementation of parallel machine learning algorithms remain a challenging task
for the project.
